# Cross-Validation

```{r echo=FALSE, message=FALSE}
rm(list = ls())

question_num <- 0
NextQ <- function() {
    question_num <<- question_num + 1
    question_num
}
```

## Discussion

**Overfitting**

From ISLR, page 32:

> When a given method yields a small training MSE but a large test MSE, we are said to be **overfitting** the data. This happens because our statistical learning procedure is working too hard to find patterns in the training data, and may be picking up some patterns that are just caused by random chance rather than by true [trends].

> When we overfit the training data, the test MSE will be very large because the supposed patterns that the method found in the training data simply don’t exist in the test data. Note that regardless of whether or not overfitting has occurred, we almost always expect the training MSE to be smaller than the test MSE because most statistical learning methods either directly or indirectly seek to minimize the training MSE. Overfitting refers specifically to the case in which a less flexible model would have yielded a smaller test MSE.

<br>
<br>

"Flexibility" in linear regression refers to the number of coefficients

- More coefficients (more predictors) = more flexible
- Fewer coefficients (fewer predictors) = less flexible

<br>
<br>
<br>
<br>
<br>

**How do we prevent overfitting?**

- Want an accuracy metric that allows us to choose which of several models will be most accurate on **new data**
    - Adjusted R-squared is a good idea but restricted to linear regression
    - Cross-validation is a much more general technique that allows estimation of the true error rate on new data (the **test error**)
- Use specific statistical learning methods suited to discourage including weak/useless predictors
    - Shrinkage methods (coming soon!)

<br>
<br>
<br>
<br>
<br>


## Exercises

**Goals**

- For what purposes would you use cross-validation?
- How is cross-validation useful for preventing overfitting?

<br>
<br>

**You can download a template RMarkdown file to start from [here](template_rmds/04-cross-validation.Rmd).**

You'll continue working on the body fat dataset from last time. The following variables were recorded.

- `fatBrozek`: Percent body fat using Brozek's equation: 457/Density - 414.2
- `fatSiri`: Percent body fat using Siri's equation: 495/Density - 450
- `density`: Density determined from underwater weighing (gm/cm^3).
- `age`: Age (years)
- `weight`: Weight (lbs)
- `height`: Height (inches)
- `neck`: Neck circumference (cm)
- `chest`: Chest circumference (cm)
- `abdomen`: Abdomen circumference (cm)
- `hip`: Hip circumference (cm)
- `thigh`: Thigh circumference (cm)
- `knee`: Knee circumference (cm)
- `ankle`: Ankle circumference (cm)
- `biceps`: Biceps (extended) circumference (cm)
- `forearm`: Forearm circumference (cm)
- `wrist`: Wrist circumference (cm)

The focus is on predicting body fat percentage using Siri's equation (`fatSiri`) from easily measured variables: age, weight, height, and circumferences.

```{r eval=FALSE}
library(ggplot2)
library(dplyr)
bodyfat_train <- read.csv("https://www.dropbox.com/s/js2gxnazybokbzh/bodyfat_train.csv?dl=1")

# Remove the fatBrozek and density variables
bodyfat_train <- bodyfat_train %>%
    select(-fatBrozek, -density)
```


`r NextQ()`. **4 models**    
    Consider the 4 models below:
    ```{r eval=FALSE}
    mod1 <- lm(fatSiri ~ age+weight+neck+abdomen+thigh+forearm, data = bodyfat_train)
    mod2 <- lm(fatSiri ~ age+weight+neck+abdomen+thigh+biceps+forearm, data = bodyfat_train)
    mod3 <- lm(fatSiri ~ age+weight+neck+chest+abdomen+hip+thigh+biceps+forearm, data = bodyfat_train)
    mod4 <- lm(fatSiri ~ ., data = bodyfat_train) # The . means all predictors
    ```
    a. Before looking at the summary tables, predict:
        - Which model will have the highest R squared?
        - Which model will have the lowest training MSE?
    b. Find/compute the R squared and MSE for the 4 models to check your answers in part a.
    c. Which model do you think will perform worst on new test data? Why?


<br>
<br>
<br>

We'll use the `caret` package to perform cross-validation (and to run many different machine learning methods throughout the course). The `caret` package is a great resource for the machine learning community because it aggregates machine learning methods written by tons of different authors in tons of different R packages into one single package. The advantage is that instead of learning 10 different styles of code for 10 different machine learning methods, we can use fairly similar code throughout the course.

Install the `caret` package by running `install.packages("caret")` in the Console.

<br>
<br>
<br>

`r NextQ()`. **Cross-validation with the `caret` package**    
    a. Use the code below to perform 10-fold cross validation for `mod1` to estimate the test MSE ($\text{CV}_{(10)}$).
        ```{r eval=FALSE}
        # Load the package
        library(caret)
        
        # Set up what type of cross-validation is desired
        train_ctrl_cv10 <- trainControl(method = "cv", number = 10)

        # To ensure the same results each time the code is run
        set.seed(253)

        # Fit (train) the model as written in mod1
        # Also supply information about the type of CV desired for evaluating the model with the trControl argument
        # na.action = na.omit prevents errors if the data has missing values
        mod1_cv10 <- train(
            fatSiri ~ age+weight+neck+abdomen+thigh+forearm,
            data = bodyfat_train,
            method = "lm",
            trControl = train_ctrl_cv10,
            na.action = na.omit
        )

        # The $ extracts components of an object
        # Peek at the "resample" part of mod1_cv10 - what info does it contain?
        mod1_cv10$resample

        # Estimate of test MSE
        # RMSE = square root of MSE
        mean(mod1_cv10$resample$RMSE^2)
        ```
    b. Adapt the code above to perform:
        - 10-fold CV for model 2
        - LOOCV for model 1
        In doing so, look carefully at the structure of the code. What parts need to be repeated? What parts don't?    
        (*Hint:* `nrow(dataset)` gives the number of cases in a dataset.)

<br>
<br>
<br>

`r NextQ()`. **Looking at the evaluation metrics**    
    A completed table of evaluation metrics is below.
    a. Which model performed the best on the training data?
    b. Which model performed best on the test set?
    c. Which model would be preferred using $\text{CV}_{(10)}$ or LOOCV estimates of the test error?
    d. How is cross-validation helping us avoid overfitting?


Model     $R^2$    Training MSE   $\text{CV}_{(10)}$    LOOCV     Test set MSE
-------- -------- -------------- -------------------- ---------- --------------
`mod1`    0.8103   14.52153       17.21062             18.16816   23.92333 
`mod2`    0.8146   14.18762       19.64114             19.29848   23.90547
`mod3`    0.816    14.08022       21.24115             20.28958   23.63958
`mod4`    0.8162   14.06917       21.88440             21.26073   24.65370


<br>
<br>
<br>

`r NextQ()`. **Practical issues: choosing $k$**    
    a. What do you think are the pros/cons of low vs. high $k$?
    b. If possible, it is advisable to choose $k$ to be a divisor of the sample size. Why do you think that is?

<br>
<br>
<br>

`r NextQ()`. **Extra! Writing R functions**    
    If you're interested in learning about writing R functions, look at the following function that the instructor used to fill out the above evaluation metrics table.
    ```{r eval=FALSE}
    bodyfat_test <- read.csv("https://www.dropbox.com/s/7gizws208u0oywq/bodyfat_test.csv?dl=1")
    evaluate_model <- function(formula) {
        train_ctrl_cv10 <- trainControl(method = "cv", number = 10)
        train_ctrl_loocv <- trainControl(method = "cv", number = nrow(bodyfat_train))
        mod_cv10 <- train(formula, data = bodyfat_train, method = "lm", trControl = train_ctrl_cv10, na.action = na.omit)
        mod_loocv <- train(formula, data = bodyfat_train, method = "lm", trControl = train_ctrl_loocv, na.action = na.omit)
        model_predictions <- predict(mod_cv10, newdata = bodyfat_test)
        test_mse <- mean((bodyfat_test$fatSiri - model_predictions)^2)
        c(
            cv10 = mean(mod_cv10$resample$RMSE^2),
            loocv = mean(mod_loocv$resample$RMSE^2),
            test = test_mse
        )
    }

    set.seed(253)
    evaluate_model(fatSiri ~ age+weight+neck+abdomen+thigh+forearm)
    evaluate_model(fatSiri ~ age+weight+neck+abdomen+thigh+biceps+forearm)
    evaluate_model(fatSiri ~ age+weight+neck+chest+abdomen+hip+thigh+biceps+forearm)
    evaluate_model(fatSiri ~ .)
    ```
    a. Step through each line and see if you can understand the structure.
    b. How would you modify the function to work on arbitrary data? How would you have to change the **function arguments** (within the parentheses on the first line)?
    c. How would you modify the function to allow the user to choose $k$?





