# Hierarchical Clustering

```{r echo=FALSE, message=FALSE}
rm(list = ls())

question_num <- 0
NextQ <- function() {
    question_num <<- question_num + 1
    question_num
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 5, fig.align = "center")
```

## Exercises

**You can download a template RMarkdown file to start from [here](template_rmds/17-hclust.Rmd).**

**These exercises are for practice and are not part of your Homework 7.**

<br>
<br>
<br>

To practice the hierarchical clustering algorithm, let's look at a small example. Suppose we collect the following sepal width and sepal length measurements from 5 irises:

```{r fig.align="center", fig.width=4, fig.height=4}
irises <- data.frame(
    Width = c(2.5,2.7,3.2,3.5,3.6),
    Length = c(5.5,6.0,4.5,5.0,4.7)
)
irises

ggplot(irises, aes(x = Width, y = Length)) +
    geom_point() + 
    geom_text(aes(label = c(1:5)), vjust = 1.5)
```

<br>
<br>
<br>

In the exercises below, you'll draw a dendrogram for these 5 specimens by hand. Sketch the following plotting frame on some scrap paper:


```{r fig.align="center", fig.width=5, fig.height=5, echo=FALSE}
ggplot(NULL, aes(y = seq(0,2.25,len=5), x = c(1:5))) +
    labs(x = "Specimen", y = "Distance")
```

<br>
<br>
<br>

`r NextQ()`.  **Step 1: First fusion**    
    - Calculate the distance between each pair of specimens:    
        ```{r eval=FALSE}
        dist(irises)
        ```    
        
        The idea:    
        
                      1       2       3       4       5
        ------- ------- ------- ------- ------- -------  
           1       0
           2               0
           3                       0
           4                               0
           5                                       0
    
    - Which pair of flowers 1-5 is most similar?
    - Draw the fusion between this pair of leaves on your plot. Be careful about the height at which you draw this fusion.

<br>
<br>
<br>

`r NextQ()`. **Step 2: Second fusion**    
    - Construct a new distance matrix reflecting the distances between each pair of branches (where 4 and 5 have now been fused). Use **complete linkage**. That is, the distance between one branch and another is the *maximum* distance between any pair of leaves in those branches. The idea:    

                      1       2       3     4 & 5
        ------- ------- ------- ------- ---------  
           1       0
           2               0
           3                       0
         4 & 5                               0
                                              

    - Which pair of branches is most similar?  Draw the fusion between this pair on the plot.

<br>
<br>
<br>

`r NextQ()`. **Step 3: Third fusion**    
    - Repeat! Construct a new distance matrix reflecting the distances between each pair of branches, those that have been fused already and those that have not.
    - Which pair of branches is most similar? Draw the fusion between this pair on the plot.

<br>
<br>
<br>

`r NextQ()`. **Step 4: Final fusion**    
    At this point, you should have 2 specimens in one cluster and 3 in another. The final step is to combine these into the tree trunk. Draw this fusion.
<br>
<br>
<br>

`r NextQ()`. Check your work in R:    
    ```{r eval=FALSE}
    iris_cluster <- hclust(dist(irises))
    plot(iris_cluster)
    ```
 
 


